Accurate evaluation of dislocation tractions

### 2020-09-21 src/wip/DanielCelisGarza/Accurate evaluation of dislocation tractions/paper/main.tex
Image forces edge perp. Add head's paper.

\documentclass[11pt]{iopart}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{booktabs}
\usepackage[noabbrev]{cleveref}
% Let us use amsmath.
\expandafter\let\csname equation*\endcsname\relax
\expandafter\let\csname endequation*\endcsname\relax
% Let us use amsmath.
\usepackage{amsmath,isomath,amssymb,bm}
\usepackage[section]{placeins}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{comment}
\usepackage[sort&compress,numbers]{natbib}
\usepackage{algorithm}
\usepackage{algpseudocode} % Package for algorithm typesetting (pseudo-code).
\usepackage{lmodern}
\graphicspath{{./images/}}
\input{macros.tex}

@@

    \address{Department of Materials, University of Oxford, Parks Road, OX1 3PH, UK}
    \begin{abstract}
        %% Text of abstract
        Dislocations generate tractions on the surface of a finite volume. Traditionally the dislocation tractions are integrated numerically over a finite element face to obtain the nodal forces; as required when using superposition to couple discrete dislocation plasticity and the finite element method. Here we implement analytic finite element nodal forces obtained using closed form expressions for the traction integrals. We compare the behaviour of the errors which arise when using numerical integration, provide insight into how and why they occur and give recommendations on avoiding numerical issues when implementing the analytic solution. Evaluating the tractions analytically in a large scale discrete dislocation plasticity simulations of a microcantilever showed a clear difference in the simulated load-displacement curve while the dislocation structure was unchanged.
    \end{abstract}

@@

&\left.

@@

\nabla\cdot{\bm{\tilde{\sigma}}} = 0\\
\bm{\tilde{\sigma}} = \bm{C}:\tilde{\bm{\varepsilon}}\\

@@

\right\}\textrm{in V}\\

@@

&\tilde{\bm{\sigma}}\cdot\bm{n} = \tilde{\bm{T}} \textrm{ on } S_T\\

@@

\tilde{\bm{u}} = \tilde{\bm{U}} \quad t>0\\ 

@@

\nabla\cdot{\bm{\hat{\sigma}}} = 0\\
\bm{\hat{\sigma}} = \bm{C}:\hat{\bm{\varepsilon}}\\

@@

&\hat{\bm{\sigma}}\cdot\bm{n} = \bm{T} - \tilde{\bm{T}} \textrm{ on } S_T\label{eq:thatbc}\\

@@

&\hat{\bm{u}} = \bm{U} - \tilde{\bm{U}} \textrm{ on } S_U \label{eq:uhatbc}

@@

\nabla\cdot\bm{\sigma}=\nabla\cdot\left(\bm{\hat{\sigma}} +\bm{\tilde{\sigma}}\right) = 0\\
\bm{\sigma} =\bm{\hat{\sigma}}+\bm{\tilde{\sigma}}= \bm{C}:\left(\hat{\bm{\varepsilon}}+\tilde{\bm{\varepsilon}}\right)=\bm{C}:{\bm{\varepsilon}}\\

@@

&\bm{\sigma}\cdot\bm{n} = \bm{T}  \textrm{ on } S_T\\

@@

\bm{u} = \bm{U} \quad t>0\\ 

@@

\begin{align}
    \label{e:shape_function}
    N^{1} &= \dfrac{1}{4}(1-s_1)(1-s_2)\\
    N^{2} &= \dfrac{1}{4}(1+s_1)(1-s_2)\nonumber\\
    N^{3} &= \dfrac{1}{4}(1+s_1)(1+s_2)\nonumber\\
    N^{4} &= \dfrac{1}{4}(1-s_1)(1+s_2)\nonumber\,.
\end{align}

@@

Equation \ref{e:ddd_fem_force} is actually a triple vector integral. It involves integrating over the surface element and along the dislocation line segment, as shown in \cref{f:force_lin_rect}. Using the isotropic Burgers vector distribution proposed in \cite{Cai2006}, the dyadic form of the stress tensor produced by a straight finite dislocation segment bounded by nodes at $\mathbf{x_1}$ and $\mathbf{x_2}$ is \cite{Queyreau},
%
\begin{align}
    \label{e:stress}
    \tns{\sigma}(\vec{x}) = & 
            - \dfrac{\mu}{8\pi} \int\limits_{\mathbf{x_1}}^{\mathbf{x_2}} \left( \dfrac{2}{R_{a}^{3}} + \dfrac{3a^2}{R_{a}^{5}} \right) \left[ \left(\vec{R} \times \vec{b}\right) \otimes \mathrm{d}\vec{x'} + \mathrm{d}\vec{x'} \otimes \left(\vec{R} \times \vec{b}\right) \right]\\
            %
            &+ \dfrac{\mu}{4\pi(1-\nu)} \int\limits_{\mathbf{x_1}}^{\mathbf{x_2}} \left( \dfrac{1}{R_{a}^{3}} + \dfrac{3a^2}{R_{a}^{5}} \right) \left[ \left(\vec{R} \times \vec{b}\right) \cdot \mathrm{d}\vec{x'} \right]\vec{I_2}\nonumber\\
            %
            &-\dfrac{\mu}{4\pi(1-\nu)} \int\limits_{\mathbf{x_1}}^{\mathbf{x_2}}  \dfrac{1}{R_{a}^{3}} \left[ \left(\vec{b} \times \mathrm{d}\vec{x'}\right) \otimes \vec{R} + \vec{R} \otimes \left(\vec{b} \times \mathrm{d}\vec{x'}\right) \right]\nonumber\\
            %
            &+ \dfrac{\mu}{4\pi(1-\nu)} \int\limits_{\mathbf{x_1}}^{\mathbf{x_2}} \dfrac{3}{R_{a}^{5}} \left[ \left(\vec{R} \times \vec{b}\right) \cdot \mathrm{d}\vec{x'} \right]\vec{R}\otimes\vec{R}\nonumber,
\end{align}
where
\begin{align}
    \vec{R} &= \vec{x} - \vec{x'} = y \vec{l} + r \vec{p} + s \vec{q}\\
    %
    R_a &= \sqrt{\vec{R} \cdot \vec{R} + a^2}\\
    %
    \mathrm{d}\vec{x'} &= -\mathrm{d} y \vec{l}
\end{align}
%
The vectors $\vec{p}$ and $\vec{q}$ are aligned with the edges of the rectangular finite element, $\vec{n} = \vec{p} \times \vec{q}$ is the element surface normal (pointing away from the dislocation), and $\vec{l}$ is parallel to the dislocation line segment as shown in figure \ref{f:force_lin_rect}. Then (provided $\vec{l}$ is not parallel to $\vec{p}$ or $\vec{q}$) $\vec{R}$ can be expressed in terms of $(\vec{l},~\vec{p},~\vec{q})$ with coefficients
%
\begin{align}
     y = \dfrac{\vec{R}\cdot \vec{n}}{\vec{l}\cdot \vec{n}} \label{se:problem},\quad
    %
     r = \dfrac{\vec{R}\cdot (\vec{q} \times \vec{l})}{\vec{p}\cdot (\vec{q} \times \vec{l})}, \quad
    %
     s = \dfrac{\vec{R}\cdot (\vec{p} \times \vec{l})}{\vec{q}\cdot (\vec{p} \times \vec{l})}\,.
\end{align}
%
Substituting (\ref{e:stress}) and (\ref{e:shape_function}) into (\ref{e:ddd_fem_force}) yields four long and messy equations (one for each FE node) that were elegantly solved by \citet{Queyreau} by utilising the fact that the triple integrals all had the form,
%
\begin{align}
    H_{ijkl} &= \int\limits_{r_{1}}^{r_{2}}\int\limits_{s_{1}}^{s_{2}}\int\limits_{y_{1}}^{y_{2}} \dfrac{r^i s^j y^k}{R_{a}^{m}}\label{e:triple_int}\\
    \textrm{when }m &= 5 \textrm{ then } i,\, j \in [0,\,3],~k \in [0,\,2]\nonumber\\
    \textrm{when }m &= 3 \textrm{ then } i,\, j \in [0,\,2],~k \in [0,\,1]\nonumber\\
    \textrm{when }m &= 1 \textrm{ then } i = j = k = 0\nonumber.
\end{align}
%
Using partial differentiation and integration by parts, they found a series of recurrence relations that lead to double and single integrals with a similar form to equation (\ref{e:triple_int}). Together they are used to construct a full solution. The recurrence relations stop working when $i = j = k = 0 \textrm{ and } m = 1,\, 3$. At which point, direct integration of the resulting single and double integrals (the last triple integrals all cancel out in the global calculation) yields six seed functions that are used as the starting point for the recurrence relations. Three of them are logarithms and three either arctangents or---if a discriminant is negative---hyperbolic arctangents. The details of the procedure can be found in \cite{Queyreau}.

@@

@@

In general one does not want a dislocation segment to intersect the surface when it is being rotated. Naively one would calculate the maximum rotational angle, $\theta_{\textrm{max}}$, to be,
%
\begin{align}
    \theta_{\textrm{max}} &= \arctan\left(\dfrac{2 d}{\left\lvert\mathbf{x_2} - \mathbf{x_1}\right\rvert}\right)\,,
\end{align}
%
where $d$ is the minimum orthogonal distance from the dislocation to the surface element (collision distance), and $\mathbf{x_1},\,\mathbf{x_2}$ are the dislocation segment node coordinates. However, $\theta_\textrm{max}$ might be too small (due to numerical precision) in cases when the segment length is too small compared to the distance to a surface element or when the segment length is much greater than $d$. Fine tuning the angle is a task that involves knowing the minimum collision distance, minimum segment length, dislocation core radius, and the compiler's implementation of mathematical functions. Given the rarity of such cases and their comparatively low impact, we chose our perturbation such that we safely avoid this problem.

@@

            \vec{x_{e,n}} &\equiv	\begin{bmatrix}

@@

    							\end{bmatrix}^{\mathsf{T}}\\

@@

        							l_{1,1} & l_{1,2} & l_{1,4} & l_{1,5}\\
        							l_{2,2} & l_{2,3} & l_{2,5} & l_{2,6}\\
        							l_{3,5} & l_{3,6} & l_{3,8} & l_{3,9}\\
        							l_{4,4} & l_{4,5} & l_{4,7} & l_{4,8}\\

@@

                            l_1\\
                            l_2\\
                            \vdots\\

@@

                        \end{bmatrix}\\

@@

        							\vec{x_{1,1}} & \vec{x_{1,2}} & \vec{x_{1,4}} & \vec{x_{1,5}}\\
        							\vec{x_{2,2}} & \vec{x_{2,3}} & \vec{x_{2,5}} & \vec{x_{2,6}}\\
        							\vec{x_{3,5}} & \vec{x_{3,6}} & \vec{x_{3,8}} & \vec{x_{3,9}}\\
        							\vec{x_{4,4}} & \vec{x_{4,5}} & \vec{x_{4,7}} & \vec{x_{4,8}}\\

@@

    								\vec{x_{1}}\\
    								\vec{x_{2}}\\
    								\vdots\\

@@

The limitations of Gauss quadrature become evident when the dislocation line segment is parallel to a surface element. \Cref{f:rel_err_par_edge} shows the relative errors for a parallel edge dislocation. In this configuration the forces due to a screw dislocations vanish as the stress field is antisymmetric.
%
\begin{figure}[t]
	\centering
    \includegraphics[width=0.8\linewidth]{par_e_z.pdf}
\caption{Log-log plot of the relative error as a function of distance from the surface element for a parallel edge dislocation (\cref{f:gauss_quad_test}b). Burgers vector, $\bm{b} = [0 1 0]$, line direction, $\bm{l} = [1 0 0]$, dislocation core and surface element parameters are the same as \cref{f:rel_err_perp_edge}. The dislocation length is fixed to $10^{6}\, b,\, x\in\left[-0.5\times10^{6},\, 0.5\times10^{6} \right]$ and bisects the surface element along the $[1 0 0]$ direction. The whole dislocation was segmented into $10^4$ pieces of length $100\, b$ to prevent the dislocation from intersecting the surface element when they were rotated to avoid the singularity. The relative error at large distances ($>10^6 b$) converges to one due to artifacting when dividing two very small double precision numbers. The absolute error always goes towards zero with increasing distance from the surface.}
\label{f:rel_err_par_edge}
\end{figure}
%
In \cref{f:rel_err_par_edge}, we observe the relative errors decrease and then converge to $\sim 1$ as the distance grows larger. The convergence to $\sim 1$ is an artifact arising from the fact that the forces on the surface element get asymptotically closer to zero with increasing distance from the surface element, which results in the division of two increasingly similar, small, non-zero numbers, resulting in $\sim1$ when done with limited precision. Given infinite precision, it is not unreasonable to assume the relative error would converge to $\sim10^{-3},~ \sim10^{-4}$. The absolute error \textit{always} decreases as the distance from the surface element increases, but absolute errors are not as useful tools of comparison. However, when the segment is close to the surface, the relative errors are large ($\gg1$) unless using a large number of quadrature points, for example \cref{f:gauss_quad_test}(a) shows that $Q\geq100$ are required when the distance $x^1_z = 15b$.

>>>>

\documentclass[11pt]{iopart}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{booktabs}
% Let us use amsmath.
\expandafter\let\csname equation*\endcsname\relax
\expandafter\let\csname endequation*\endcsname\relax
% Let us use amsmath.
\usepackage{amsmath,isomath,amssymb,bm}
\usepackage[noabbrev]{cleveref}
\usepackage[section]{placeins}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{comment}
\usepackage[sort&compress,numbers]{natbib}
\usepackage{algorithm}
\usepackage{algpseudocode} % Package for algorithm typesetting (pseudo-code).
\usepackage{lmodern}
\graphicspath{{./images/}}
\input{macros.tex}

@@

\address{Department of Materials, University of Oxford, Parks Road, OX1 3PH, UK}

\begin{abstract}
    %% Text of abstract
    Dislocations generate tractions on the surface of a finite volume. Traditionally the dislocation tractions are integrated numerically over a finite element face to obtain the nodal forces; as required when using superposition to couple discrete dislocation plasticity and the finite element method. Here we implement analytic finite element nodal forces obtained using closed form expressions for the traction integrals. We compare the behaviour of the errors which arise when using numerical integration, provide insight into how and why they occur and give recommendations on avoiding numerical issues when implementing the analytic solution. Evaluating the tractions analytically in a large scale discrete dislocation plasticity simulations of a microcantilever showed a clear difference in the simulated load-displacement curve while the dislocation structure was unchanged.
\end{abstract}

@@

     & \left.

@@

        \nabla\cdot{\bm{\tilde{\sigma}}} = 0                  \\
        \bm{\tilde{\sigma}} = \bm{C}:\tilde{\bm{\varepsilon}} \\

@@

    \right\}\textrm{in V}                                                \\

@@

     & \tilde{\bm{\sigma}}\cdot\bm{n} = \tilde{\bm{T}} \textrm{ on } S_T \\

@@

        \tilde{\bm{u}} = \tilde{\bm{U}} \quad t>0 \\

@@

        \nabla\cdot{\bm{\hat{\sigma}}} = 0                \\
        \bm{\hat{\sigma}} = \bm{C}:\hat{\bm{\varepsilon}} \\

@@

     & \hat{\bm{\sigma}}\cdot\bm{n} = \bm{T} - \tilde{\bm{T}} \textrm{ on } S_T\label{eq:thatbc} \\

@@

     & \hat{\bm{u}} = \bm{U} - \tilde{\bm{U}} \textrm{ on } S_U \label{eq:uhatbc}

@@

        \nabla\cdot\bm{\sigma}=\nabla\cdot\left(\bm{\hat{\sigma}} +\bm{\tilde{\sigma}}\right) = 0                                                         \\
        \bm{\sigma} =\bm{\hat{\sigma}}+\bm{\tilde{\sigma}}= \bm{C}:\left(\hat{\bm{\varepsilon}}+\tilde{\bm{\varepsilon}}\right)=\bm{C}:{\bm{\varepsilon}} \\

@@

     & \bm{\sigma}\cdot\bm{n} = \bm{T}  \textrm{ on } S_T \\

@@

        \bm{u} = \bm{U} \quad t>0 \\

@@

\begin{align}
    \label{e:shape_function}
    N^{1} & = \dfrac{1}{4}(1-s_1)(1-s_2)             \\
    N^{2} & = \dfrac{1}{4}(1+s_1)(1-s_2)\nonumber    \\
    N^{3} & = \dfrac{1}{4}(1+s_1)(1+s_2)\nonumber    \\
    N^{4} & = \dfrac{1}{4}(1-s_1)(1+s_2)\nonumber\,.
\end{align}

@@

Equation \ref{e:ddd_fem_force} is actually a triple vector integral. It involves integrating over the surface element and along the dislocation line segment, as shown in \cref{f:force_lin_rect}. Using the isotropic Burgers vector distribution proposed in \cite{Cai2006}, the dyadic form of the stress tensor produced by a straight finite dislocation segment bounded by nodes at $\mathbf{x_1}$ and $\mathbf{x_2}$ is \cite{Queyreau},
    \label{e:stress}
    \tns{\sigma}(\vec{x}) = &
    - \dfrac{\mu}{8\pi} \int\limits_{\mathbf{x_1}}^{\mathbf{x_2}} \left( \dfrac{2}{R_{a}^{3}} + \dfrac{3a^2}{R_{a}^{5}} \right) \left[ \left(\vec{R} \times \vec{b}\right) \otimes \mathrm{d}\vec{x'} + \mathrm{d}\vec{x'} \otimes \left(\vec{R} \times \vec{b}\right) \right]  \\
                            & + \dfrac{\mu}{4\pi(1-\nu)} \int\limits_{\mathbf{x_1}}^{\mathbf{x_2}} \left( \dfrac{1}{R_{a}^{3}} + \dfrac{3a^2}{R_{a}^{5}} \right) \left[ \left(\vec{R} \times \vec{b}\right) \cdot \mathrm{d}\vec{x'} \right]\vec{I_2}\nonumber                  \\
                            & -\dfrac{\mu}{4\pi(1-\nu)} \int\limits_{\mathbf{x_1}}^{\mathbf{x_2}}  \dfrac{1}{R_{a}^{3}} \left[ \left(\vec{b} \times \mathrm{d}\vec{x'}\right) \otimes \vec{R} + \vec{R} \otimes \left(\vec{b} \times \mathrm{d}\vec{x'}\right) \right]\nonumber \\
                            & + \dfrac{\mu}{4\pi(1-\nu)} \int\limits_{\mathbf{x_1}}^{\mathbf{x_2}} \dfrac{3}{R_{a}^{5}} \left[ \left(\vec{R} \times \vec{b}\right) \cdot \mathrm{d}\vec{x'} \right]\vec{R}\otimes\vec{R}\nonumber,
where
    \vec{R}            & = \vec{x} - \vec{x'} = y \vec{l} + r \vec{p} + s \vec{q} \\
    R_a                & = \sqrt{\vec{R} \cdot \vec{R} + a^2}                     \\
    \mathrm{d}\vec{x'} & = -\mathrm{d} y \vec{l}
The vectors $\vec{p}$ and $\vec{q}$ are aligned with the edges of the rectangular finite element, $\vec{n} = \vec{p} \times \vec{q}$ is the element surface normal (pointing away from the dislocation), and $\vec{l}$ is parallel to the dislocation line segment as shown in figure \ref{f:force_lin_rect}. Then (provided $\vec{l}$ is not parallel to $\vec{p}$ or $\vec{q}$) $\vec{R}$ can be expressed in terms of $(\vec{l},~\vec{p},~\vec{q})$ with coefficients
    y = \dfrac{\vec{R}\cdot \vec{n}}{\vec{l}\cdot \vec{n}} \label{se:problem},\quad
    r = \dfrac{\vec{R}\cdot (\vec{q} \times \vec{l})}{\vec{p}\cdot (\vec{q} \times \vec{l})}, \quad
    s = \dfrac{\vec{R}\cdot (\vec{p} \times \vec{l})}{\vec{q}\cdot (\vec{p} \times \vec{l})}\,.
Substituting (\ref{e:stress}) and (\ref{e:shape_function}) into (\ref{e:ddd_fem_force}) yields four long and messy equations (one for each FE node) that were elegantly solved by \citet{Queyreau} by utilising the fact that the triple integrals all had the form,
    H_{ijkl}        & = \int\limits_{r_{1}}^{r_{2}}\int\limits_{s_{1}}^{s_{2}}\int\limits_{y_{1}}^{y_{2}} \dfrac{r^i s^j y^k}{R_{a}^{m}}\label{e:triple_int} \\
    \textrm{when }m & = 5 \textrm{ then } i,\, j \in [0,\,3],~k \in [0,\,2]\nonumber                                                                         \\
    \textrm{when }m & = 3 \textrm{ then } i,\, j \in [0,\,2],~k \in [0,\,1]\nonumber                                                                         \\
    \textrm{when }m & = 1 \textrm{ then } i = j = k = 0\nonumber.
Using partial differentiation and integration by parts, they found a series of recurrence relations that lead to double and single integrals with a similar form to equation (\ref{e:triple_int}). Together they are used to construct a full solution. The recurrence relations stop working when $i = j = k = 0 \textrm{ and } m = 1,\, 3$. At which point, direct integration of the resulting single and double integrals (the last triple integrals all cancel out in the global calculation) yields six seed functions that are used as the starting point for the recurrence relations. Three of them are logarithms and three either arctangents or---if a discriminant is negative---hyperbolic arctangents. The details of the procedure can be found in \cite{Queyreau}.

@@

In simulations, the tractions are manifested in the reaction forces calculated by the FE solver at FE nodes. In order to validate and compare the differences in tractions have on dislocation plasticity simulations we compare the analytic expressions for infinite dislocations in inhomogenous media by \citet{head1953edge}. The original paper has a few typos that have been replicated in other sources so we write the correct expressions for an edge dislocation parallel to a surface described by the line $x = 0$, where the dislocation coordinates are $(a,~c)$ and we are looking at the $xy$-plane where $\vec{b}$ is perpendicular to the surface in \cref{e:imageStressAnalyticEdge1} and parallel to the surface in \cref{e:imageStressAnalyticEdge2}.
\begin{subequations}
    \begin{align}\label{e:imageStressAnalyticEdge1}
        \sigma_{xx} = D (y - c) & \left\{-\dfrac{3 (x - a)^2 + (y - c)^2}{[(x - a)^2 + (y - c)^2]^2}\right.                                                                                                        \\\nonumber
                                & \left. +\dfrac{3 (x + a)^2 + (y - c)^2}{[(x + a)^2 + (y - c)^2]^2} + 4 a x \dfrac{3 (x + a)^2 - (y - c)^2}{[(x + a)^2 + (y - c)^2]^3}\right\},                                   \\
        \sigma_{yy} = D (y - c) & \left\{\dfrac{(x - a)^2 - (y - c)^2}{[(x - a)^2 + (y - c)^2]^2}                                                 \right.                                                          \\\nonumber
                                & \left. -\dfrac{(x + a)^2 - (y - c)^2}{[(x + a)^2 + (y - c)^2]^2} + 4 a (2 a - x) \dfrac{(x + a)^2 + (3 x + 2 a) (y - c)^2}{[(x + a)^2 + (y - c)^2]^3}\right\},                   \\
        \sigma_{xy} = D         & \left\{(x - a) \dfrac{(x - a)^2 - (y - c)^2}{[(x - a)^2 + (y - c)^2]^2} \right.                                                                                                  \\\nonumber
                                & \left. -(x + a) \dfrac{(x + a)^2 - (y - c)^2}{[(x + a)^2 + (y - c)^2]^2} + 2 a \dfrac{6 x (x + a) (y - c)^2 - (x - a) (x + a)^3 - (y - c)^4}{[(x + a)^2 + (y - c)^2]^3}\right\},
    \end{align}
\end{subequations}

\begin{subequations}
    \begin{align}\label{e:imageStressAnalyticEdge2}
        \sigma_{xx} = D (y - c) & \left\{-\dfrac{3 (x - a)^2 + (y - c)^2}{[(x - a)^2 + (y - c)^2]^2}\right.                                                                                                        \\\nonumber
                                & \left. +\dfrac{3 (x + a)^2 + (y - c)^2}{[(x + a)^2 + (y - c)^2]^2} + 4 a x \dfrac{3 (x + a)^2 - (y - c)^2}{[(x + a)^2 + (y - c)^2]^3}\right\},                                   \\
        \sigma_{yy} = D (y - c) & \left\{\dfrac{(x - a)^2 - (y - c)^2}{[(x - a)^2 + (y - c)^2]^2}                                                 \right.                                                          \\\nonumber
                                & \left. -\dfrac{(x + a)^2 - (y - c)^2}{[(x + a)^2 + (y - c)^2]^2} + 4 a (2 a - x) \dfrac{(x + a)^2 + (3 x + 2 a) (y - c)^2}{[(x + a)^2 + (y - c)^2]^3}\right\},                   \\
        \sigma_{xy} = D         & \left\{(x - a) \dfrac{(x - a)^2 - (y - c)^2}{[(x - a)^2 + (y - c)^2]^2} \right.                                                                                                  \\\nonumber
                                & \left. -(x + a) \dfrac{(x + a)^2 - (y - c)^2}{[(x + a)^2 + (y - c)^2]^2} + 2 a \dfrac{6 x (x + a) (y - c)^2 - (x - a) (x + a)^3 - (y - c)^4}{[(x + a)^2 + (y - c)^2]^3}\right\},
    \end{align}
\end{subequations}

@@

In general one does not want a dislocation segment to intersect the surface when it is being rotated. Naively one would calculate the maximum rotational angle, $\theta_{\textrm{max}}$, to be,
%
\begin{align}
    \theta_{\textrm{max}} & = \arctan\left(\dfrac{2 d}{\left\lvert\mathbf{x_2} - \mathbf{x_1}\right\rvert}\right)\,,
\end{align}
%
where $d$ is the minimum orthogonal distance from the dislocation to the surface element (collision distance), and $\mathbf{x_1},\,\mathbf{x_2}$ are the dislocation segment node coordinates. However, $\theta_\textrm{max}$ might be too small (due to numerical precision) in cases when the segment length is too small compared to the distance to a surface element or when the segment length is much greater than $d$. Fine tuning the angle is a task that involves knowing the minimum collision distance, minimum segment length, dislocation core radius, and the compiler's implementation of mathematical functions. Given the rarity of such cases and their comparatively low impact, we chose our perturbation such that we safely avoid this problem.

@@

        \vec{x_{e,n}} & \equiv	\begin{bmatrix}

@@

        \end{bmatrix}^{\mathsf{T}}            \\

@@

                l_{1,1} & l_{1,2} & l_{1,4} & l_{1,5} \\
                l_{2,2} & l_{2,3} & l_{2,5} & l_{2,6} \\
                l_{3,5} & l_{3,6} & l_{3,8} & l_{3,9} \\
                l_{4,4} & l_{4,5} & l_{4,7} & l_{4,8} \\

@@

            l_1    \\
            l_2    \\
            \vdots \\

@@

        \end{bmatrix}                          \\

@@

                \vec{x_{1,1}} & \vec{x_{1,2}} & \vec{x_{1,4}} & \vec{x_{1,5}} \\
                \vec{x_{2,2}} & \vec{x_{2,3}} & \vec{x_{2,5}} & \vec{x_{2,6}} \\
                \vec{x_{3,5}} & \vec{x_{3,6}} & \vec{x_{3,8}} & \vec{x_{3,9}} \\
                \vec{x_{4,4}} & \vec{x_{4,5}} & \vec{x_{4,7}} & \vec{x_{4,8}} \\

@@

            \vec{x_{1}} \\
            \vec{x_{2}} \\
            \vdots      \\

@@

The limitations of Gauss quadrature become evident when the dislocation line segment is parallel to a surface element. \Cref{f:rel_err_par_edge} shows the relative errors for a parallel edge dislocation. In this configuration the forces due to a screw dislocations vanish as the stress field is antisymmetric.
%
\begin{figure}[t]
    \centering
    \includegraphics[width=0.8\linewidth]{par_e_z.pdf}
    \caption{Log-log plot of the relative error as a function of distance from the surface element for a parallel edge dislocation (\cref{f:gauss_quad_test}b). Burgers vector, $\bm{b} = [0 1 0]$, line direction, $\bm{l} = [1 0 0]$, dislocation core and surface element parameters are the same as \cref{f:rel_err_perp_edge}. The dislocation length is fixed to $10^{6}\, b,\, x\in\left[-0.5\times10^{6},\, 0.5\times10^{6} \right]$ and bisects the surface element along the $[1 0 0]$ direction. The whole dislocation was segmented into $10^4$ pieces of length $100\, b$ to prevent the dislocation from intersecting the surface element when they were rotated to avoid the singularity. The relative error at large distances ($>10^6 b$) converges to one due to artifacting when dividing two very small double precision numbers. The absolute error always goes towards zero with increasing distance from the surface.}
    \label{f:rel_err_par_edge}
\end{figure}
%
In \cref{f:rel_err_par_edge}, we observe the relative errors decrease and then converge to $\sim 1$ as the distance grows larger. The convergence to $\sim 1$ is an artifact arising from the fact that the forces on the surface element get asymptotically closer to zero with increasing distance from the surface element, which results in the division of two increasingly similar, small, non-zero numbers, resulting in $\sim 1$ when done with limited precision. Given infinite precision, it is not unreasonable to assume the relative error would converge to $\sim10^{-3},~ \sim10^{-4}$. The absolute error \textit{always} decreases as the distance from the surface element increases, but absolute errors are not as useful tools of comparison. However, when the segment is close to the surface, the relative errors are large ($\gg1$) unless using a large number of quadrature points, for example \cref{f:gauss_quad_test}(a) shows that $Q\geq100$ are required when the distance $x^1_z = 15b$.
